The Talking Heads experiment

### 2015-04-29 chap11/chap11.tex
part 2 of martin haspelmath's comments

While introducing these experiments I will highlight two conceptual breakthroughs: semiotic networks and language strategies. 
\begin{itemize}
\item A {\itshape semiotic network} consists of a weighted network that links words to concepts, concepts to prototypes, and prototypes to 
sensory experiences. Each agent constructs his own semiotic network but through consecutive 
interactions these networks become progressively similar. From this perspective, setting up a language game requires a definition and 
operationalisation how new nodes are created in this network, how new links are made between 
the nodes, and how the weights between nodes are adjusted. 
\item A {\itshape language strategy} is helpful to think about how we can go beyond individual words and towards 
more complex grammatical expressions. It was already introduced in the previous chapter. 
A language strategy is a particular way to deal syntactically and semantically 
with one domain of meaning. It includes ways for conceptualising, expressing, acquiring and expanding a specific domain. 
For example, the colour domain requires not only strategies for basic colour terms (like \smplenquote{blue}), but also for 
graded membership expressions (\smplenquote{very blue}), composite colours, (\smplenquote{greenish blue}), analogical colours (\smplenquote{sky blue}), a.o. 
The Talking Heads experiments used a single strategy that was entirely hard-coded: discrimination trees for categorisation 
and words without syntax to express distinctive categories. Research efforts moved towards the study of other strategies, and 
to the question how such strategies may arise (discussed in the next chapter). 
\end{itemize}
The remainder of the present chapter illustrates these two notions through different language game experiments: The Proper 
Naming Game and the Action Naming Game illustrate the use of semiotic networks. The Colour Description Game and the 
Spatial Language Game illustrate the use of language strategies. These experiments also show how the focus of research 
progressively moved towards the emergence of grammar in co-evolution with more complex compositional, but still 
grounded, semantics.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap10/figs/png-perception.pdf}}
\caption{\label{fig:png-perception}A robot is looking at consecutive scenes (from top to bottom) and tracking the objects in the scene. One object (labelled $O_{716}$)
is moved by a human experimenter.  We see from left to right: the original source image, results of region detection and 
colour segmentation, the (average) value for one dimension of one object ($O_{716}$), the segmented and labelled objects, 
and the world model, including the position of the other robot (indicated by a pointing arrow). }
\end{figure}
\clearpage
We call the feature vector for each object a {\itshape sensory experience}. \is{sensory experience}
It is represented using a spiking 
diagram, with a spike for each dimension and the value on that dimension marked (see 
\figref{fig:prototypes-identifiers-words}, left-most column). Real
world vision is difficult enough, but because the robots look at the
scene from different points of view, they often do not carve out the
same regions and they will certainly see different views of the
objects and locate them in different positions within their own
egocentric reference frame. So sensory experiences of the same scene
are always going to be different. This makes coordination and hence successful
communication very challenging.

@@

\subsection{Semiotic networks}
We have already seen from earlier experiments that the solution to these various issues lies in setting up
the right kind of \dblenquote{semiotic dynamics}. In the present case, this dynamics should gradually coordinate
sensory experiences, prototypical views, individuals, 
and names, both within a single individual and across the
population. This led to the notion of a {\itshape semiotic network}, which was a significant conceptual advance 
in the methodology of conceiving and carrying out language game experiments.

@@

1. The robots need a {\itshape repertoire of actions}. So they need a sensori-motor system that allows them to carry out 
repeatedly and reliably an action. They can themselves observe the action both in the visual domain,  
looking at their own body or looking at their body in the mirror, or through proprioception, using sensors attached 
to the motors and joints on the body. There are many ways in which an action repertoire can be built up. One way 
is through motor babbling: The robot makes random movements and thus explores the sensori-motor 
space in order to find possible coherent actions. In the experiments reported here, the motor actions are acquired 
by kinesthetic learning: The experimenter moves parts of the body of the robot. These movements are recorded by the robot
and he can then play the movement back.

@@

In our group, Tony Belpaeme became the main ph.D researcher for Colour Naming, finishing his thesis 
in 2005.\footnote{The first extensive paper on this research appeared as: \cite{Steels:2005}.}
This work was then continued by Joris Bleys, who managed to scale up from colour naming to colour description with 
a thesis defended in 2010.\footnote{This thesis is published as another volume in the present series as: \cite{Bleys:2014}}
The colour domain has attracted several other language game researchers, attempting to explain 
universal trends in colour categorisation.\footnote{Most notably: \cite{Puglisi:2008}, \cite{Baronchelli:2010}.}

@@

When the \textsc{qrio} humanoid robots became available, we decided to use this existing work for pushing forward the state of 
the art in colour language research in two ways:
\begin{itemize}
\item To use more realistic circumstances of object perception. Because 
robots are moving around in three dimensions and see the situation from different points of view, the reflection 
of light changes and is different for speaker and hearer (see \tabref{f:scene-example}). This difference did not 
play a role in earlier experiments but 
it certainly plays an important role in the stabilisation and choice of colour prototypes. 
\item To focus on colour descriptions beyond basic colour terms, i.e. descriptions like 
\smplenquote{slightly blue}, \smplenquote{blue-green}, \smplenquote{bluish green}, \smplenquote{very bright}, \smplenquote{shiny yellow}, \smplenquote{dark red}, \smplenquote{sky blue}, 
\smplenquote{very dark bluish red}, etc. Indeed, it turns out that 
basic colour terms are used only in about 10 to 15 \% of the occurrences of colour descriptions. The rest are 
more complex expressions. Complex colour descriptions require scaling up the semantics and introducing grammar, they 
also require the introduction of language strategies as a layer of abstraction 
above lexical and grammatical constructions. 
\end{itemize}

@@

A language strategy has two components: (i) a particular way of conceptualising reality (and learning the categories involved
in this conceptualisation) and (ii) a particular grammatical pattern that suggests to the listener what strategy is intended 
and provides the contents to execute the strategy. For example, in the phrase \smplenquote{very green}, the combination of an adverb 
\smplenquote{very} and a basic colour adjective \smplenquote{green} suggests a \dblenquote{graded membership strategy} where not only a basic colour prototype
(\smplenquote{green}) is introduced but also how close the sample is to this prototype (\smplenquote{very}). I will now introduce first how 
the conceptual aspect of a language strategy is defined using a procedural semantics system called IRL and then how the 
linguistic aspect is defined, using Fluid Construction Grammar.

@@

To work out the procedural semantics hypothesis, we need to address the question what the primitives are of 
these semantic programs, and what computational formalism is most suited to combine them. Already in 2000, I reported
a representation system for computational semantics, called IRL.\footnote{
The earliest paper introducing IRL is: \cite{Steels:2000}. 
A more recent description of a new implementation is reported as: \cite{Spranger:2012}.}
IRL stands for Incremental Recruitment Language, for reasons that will become clear soon. \is{IRL}
IRL uses a constraint language approach. Constraint programming languages were pioneered in the  \is{Constraint Programming}
seventies by Borning, Steele, and others. It is a very active field with its own conferences, 
organisations and journals (see \url{http://www.informatik.uni-trier.de/ley/db/journals/constraints/}) 
with wide applications in user interfaces, scheduling problems, design, games, etc.

@@

Constraint networks are an obvious choice for grounded (natural) language processing. \is{constraint network}
For example, to interpret the sentence `the ball behind
the block', it is possible that there are many blocks, many balls, but only one ball that is behind a block. So the 
unique referent of \smplenquote{the ball}, \smplenquote{the block} and of the total noun phrase only become clear when 
all constraints have been applied. Natural languages clearly do not supply control information but assume that 
the interpreter is smart enough to interpret constraints based on data flow. The other properties of constraint 
programming are also important. The constraint network planner can be used for conceptualising what to say and 
the constraint network matcher is highly useful in parsing an incomplete language fragment or in learning the 
meaning of a new grammatical construction.

@@

What is the nature of the constraints that we will need for the semantics of natural language? 
Take a phrase like \smplenquote{the ball behind the block}. It invokes the following constraints, also called 
semantic operations: 
\begin{itemize}
\item \smplenquote{the ball} evokes a constraint to use a prototype [\textsc{ball}] to find a referent in the current scene.
\item The same holds for \smplenquote{the block} which uses another prototype [\textsc{block}]. 
\item The preposition \smplenquote{behind} implies another constraint, namely to use the spatial relation [\textsc{behind}] to filter a 
set of objects (namely those found by using the [\textsc{ball}] prototype) and retain only those that satisfy the behind-relation 
to an object in another set. 
\end{itemize}

@@

So IRL is in fact a way to integrate the broad set of mechanisms used in 
pattern recognition (including neural networks) and AI (including machine learning). 
For example, an operation could implement a radial-basis-function network for doing categorisation and 
use the network to identify an object in a set. It could not only use the network but also adapt 
the weights in the network. Another operation could implement categorisation based 
on discrimination trees (as discussed in \chapref{chap:4}) and would not only do it, but also expand the discrimination trees 
when necessary. Other operations could implement set-theoretic operations, 
for the semantics of connectives like {\itshape and}, {\itshape or}, and {\itshape not}, by carrying out
{\itshape union}, {\itshape intersection} and {\itshape complement} operations over sets. There is no 
limit to the possible semantic operations, except that they have to terminate in finite time, possibly giving up when 
they cannot reach a valid result. IRL just provides a framework to define these operations and make them available 
to be recruited by the constraint network planner. IRL uses LISP as the host language, which means that all 
the computational machinery of LISP is available to implement semantic operations and associated learning algorithms.

@@

\enlargethispage{1\baselineskip}
1. The first strategy is the {\bfshape Basic Colour Strategy} which implements a prototype-based approach to 
the basic colour categories. This approach is similar to the one used in the Proper Naming Game, except that 
only the three colour-dimensions are profiled (the opponent channels red-green, yellow-blue and the brightness channel). 
The strategy is similar to the one shown earlier. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour-lenient ?s3 ?s2 ?cc)
(select-most-activated ?t ?s3)
\end{verbatim}
When the speaker says \smplenquote{green}, this network is used together with a specific binding of ?cc to the prototype for 
green (denoted as [green]). 
\begin{verbatim}
(bind colour-category ?cc [green])
\end{verbatim}
A graphical representation of this network is shown in \figref{fig:basic-strat}.

@@

2. The second strategy is the {\bfshape Graded Membership Strategy} which categorises the topic first using 
a basic colour but then qualifies how well its colour fits with the prototype. For example, we can say \smplenquote{greenish} or
\smplenquote{very green}. Some languages (such as Tarahumara spoken in Northern Mexico) obligatory mark graded membership using 
suffixes. The core of the constraint network is defined as follows. A graphical representation of the network is shown
in \figref{fig:graded}. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour ?s3 ?s2 ?cc)
(filter-by-membership ?s4 ?s3 ?mc)
(select-most-activated ?t ?s4)
\end{verbatim}
This network grabs the objects in the context and binds them to ?s1 using the equal-to-context operation. 
The profile-colour-dimensions operation picks out the colour properties of the various objects in ?s1 resulting in 
a new set ?s2. The filter-by-colour operation computes for each object in the set ?s2
how well they match with the prototype bound to ?cc, constructing a new set ?s3. The filter-by-membership operation 
then computes how well 
each object in ?s3 satisfies the graded membership qualification bound to ?mc, 
constructing ?s4. And then the select-most-activated operation picks out the best candidate and binds it to ?t. 
Other data flows within the same network are possible, for example to find the bindings of ?mc and ?cc given 
a context and a referent ?t that needs to be identified.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/graded.pdf}}
\caption{\label{fig:graded}Constraint network implementing the Graded Membership Strategy. It builds further on the network for Basic Colour
Categorisation but adds a constraint to handle graded membership. }
\end{figure}
\enlargethispage{1\baselineskip}
The network is entirely general for all expressions that use graded membership and gets instantiated for a particular 
case by specifying the semantic entities involved using the bind operation. For example, Tarahumara has the word 
\smplenquote{sita} for a particular colour prototype and \smplenquote{kame} for one type of graded membership: 
\begin{verbatim}
(bind colour-category ?cc [sita])
(bind membership-category ?mc [kame]) 
\end{verbatim}

3. The third strategy is the {\bfshape Category Combination Strategy} which compounds a category into two new ones, in 
expressions like \smplenquote{greenish blue} or \smplenquote{red orange}. This compounding is not a simple intersection. The set of objects  
that are \smplenquote{greenish blue} is not found by taking all the ones that are green and all the ones that are blue and then 
the intersection of the two. Rather, it appears that the discretisation of the colour landscape is scaled with respect 
to the colour which appears as head of the phrase and then categorisation is done. This is achieved with the IRL network 
shown in \figref{fig:combi}. 
%http://www.workwithcolour.com/red-orange-colour-hue-range-01.htm

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/combi.pdf}}
\caption{\label{fig:combi}Constraint network implementing the Colour Combination strategy leading to expressions like \smplenquote{greenish blue}. 
The filter-by-colour operation now has an extra argument, namely category-sets,  
and there is a draw-category-set-to-category operation that rescales the basic colour category set.}
\end{figure}

4. The next strategy is the {\bfshape Basic Modification Strategy} which modifies a colour along the brightness dimension, as in 
\smplenquote{light green}, \smplenquote{dull yellow}, etc. It is implemented by projecting the modifying dimension onto the basic colour space 
whereby the colour being modified lies on the projecting vector and a new discretisation of the colour space can be 
performed. The network for this strategy is shown in \figref{fig:modi}. 
%http://www.workwithcolour.com/red-orange-colour-hue-range-01.htm

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/modi.pdf}}
\caption{\label{fig:modi}Constraint network for the basic modification strategy, for expressions like \smplenquote{light green}.}
\end{figure}

@@

1. There are first {\itshape lexical constructions} that associate a word with the introduction of a semantic entity  
such as a colour category or a qualifier for indicating graded membership. So the meaning for the word \smplenquote{yellow} 
introduces a \textsc{bind}-operation, as in: 
\begin{verbatim}
((bind colour-category ?c [yellow]))
\end{verbatim}
When the word \smplenquote{yellow} is encountered, a unit for this word is created with this \textsc{bind}-operation as meaning and the 
word-string itself as the form. It is the unit called Yellow-unit in \figref{fig:syntactic-graded}. 
The construction also introduces a lexical category (part of speech) for the word (in this case the lexical category 
is colour-category) and specifies which variables in the meaning can be linked with other variables in other
units. The variables are typed. In this case there is one variable ?c which is of type colour-category.

@@

2. Next there are {\itshape functional constructions} that determine in which semantic operation the meaning of a word 
will play a role. It is necessary to have this layer above lexical categories because the same word/category can have many 
different functions. The choice is determined by the lexical category, and the syntactic and semantic context as 
defined in functional and phrasal constructions. For example, an adjective can appear both in an adjectival 
function (as in \smplenquote{the red block}) and as a predicate (as in \smplenquote{the block is red}) but it can also be coerced 
into a nominal function (as in \smplenquote{red is a beautiful colour}). The same word can also belong to more than one 
lexical category, for example, \smplenquote{walk} can be a verb (\smplenquote{He walks home}) as well as a noun (\smplenquote{I take a walk}). \is{functional 
constructions}

@@

3. Finally there are {\itshape phrasal constructions} that pull together the functional units in order to create a phrase. 
\is{phrasal construction}
The phrase usually has additional semantic operations as well. \figref{fig:syntactic-graded},left shows an example. 
The ColourPhrase-Unit is constructed on-top of the Colour-Unit. It adds three additional operations to the meaning: 
\begin{verbatim}
((equal-to-context ?s1)
 (profile-colour-dimensions ?s2 ?s1)
 (select-most-activated ?t ?s3))
\end{verbatim}
It links to the variables supplied by Colour-unit (namely ?s3 and ?s2) and the unit itself has a variable that can 
be linked further, namely ?t, which is the topic of the whole utterance. 
The constituent is called Colour-Phrase.

\figref{fig:syntactic-graded}, right shows another example, namely the structure for an instantiation of the 
graded-membership strategy for the Tarahumara utterance \smplenquote{sita kame}. It is again a Colour-Phrase.
We see that there are now 4 levels of hierarchy. The ColourMembership-Unit pulls together a colour-category and 
a membership-category to form a new colour-category which then becomes a ColourPhrase-Unit again.
\subsection{Influence of embodiment}
% {\bfshape Influence of embodiment}\\

@@

I report here just briefly one additional experiment, relevant for understanding 
the impact of embodiment on language. The experiment compares three conditions (see \figref{fig:grounding-comparison}): 
(i) simulated perception (left column), (ii) shared grounded perception (middle column), meaning that both robots 
are given the same perception, and (iii) individual grounded 
perception (right column). The latter is the normal case when two robots are looking at the scene. 
It furthermore compares results for the Basic Naming Strategy (implemented with the IRL constraint network in 
\figref{fig:modi}) for four experimental conditions: 
\begin{enumerate}
\item There is a {\itshape baseline} experiment which starts out with the basic English colour categories provided 
by design to the agents. 
\item An {\itshape acquisition} experiment, where one agent (the tutor) is initialised with a set of basic colour categories and 
the other agent (the learning) has to acquire them. 
\item A {\itshape formation} experiment which involves a population of ten agents which develop a 
colour ontology and lexicon from scratch. 
\item And an {\itshape adaptation} experiment that starts with agents that are initialised with the 
English categories but then are allowed to further adapt their lexicons depending on the challenges of the situations 
they encounter. 
\end{enumerate}

>>>>

While introducing these experiments I will highlight two conceptual breakthroughs: semiotic networks and language strategies. 
\begin{itemize}
\item A {\scshape semiotic network} consists of a weighted network that links words to concepts, concepts to prototypes, and prototypes to 
sensory experiences. Each agent constructs his own semiotic network but through consecutive 
interactions these networks become progressively similar. From this perspective, setting up a language game requires a definition and 
operationalisation how new nodes are created in this network, how new links are made between 
the nodes, and how the weights between nodes are adjusted. 
\item A {\scshape language strategy} is helpful to think about how we can go beyond individual words and towards 
more complex grammatical expressions. It was already introduced in the previous chapter. 
A language strategy is a particular way to deal syntactically and semantically 
with one domain of meaning. It includes ways for conceptualising, expressing, acquiring and expanding a specific domain. 
For example, the colour domain requires not only strategies for basic colour terms (like \smplenquote{blue}), but also for 
graded membership expressions (\smplenquote{very blue}), composite colours, (\smplenquote{greenish blue}), analogical colours (\smplenquote{sky blue}), a.o. 
The Talking Heads experiments used a single strategy that was entirely hard-coded: discrimination trees for categorisation 
and words without syntax to express distinctive categories. Research efforts moved towards the study of other strategies, and 
to the question how such strategies may arise (discussed in the next chapter). 
\end{itemize}
The remainder of the present chapter illustrates these two notions through different language game experiments: The Proper 
Naming Game and the Action Naming Game illustrate the use of semiotic networks. The Colour Description Game and the 
Spatial Language Game illustrate the use of language strategies. These experiments also show how the focus of research 
progressively moved towards the emergence of grammar in co-evolution with more complex compositional, but still 
grounded, semantics.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap10/figs/png-perception.pdf}}
\caption{\label{fig:png-perception}A robot is looking at consecutive scenes (from top to bottom) and tracking the objects in the scene. One object (labelled $O_{716}$)
is moved by a human experimenter.  We see from left to right: the original source image, results of region detection and 
colour segmentation, the (average) value for one dimension of one object ($O_{716}$), the segmented and labelled objects, 
and the world model, including the position of the other robot (indicated by a pointing arrow). }
\end{figure}
\clearpage
We call the feature vector for each object a {\scshape sensory experience}. \is{sensory experience}
It is represented using a spiking 
diagram, with a spike for each dimension and the value on that dimension marked (see 
\figref{fig:prototypes-identifiers-words}, left-most column). Real
world vision is difficult enough, but because the robots look at the
scene from different points of view, they often do not carve out the
same regions and they will certainly see different views of the
objects and locate them in different positions within their own
egocentric reference frame. So sensory experiences of the same scene
are always going to be different. This makes coordination and hence successful
communication very challenging.

@@

\subsection{Semiotic networks}
We have already seen from earlier experiments that the solution to these various issues lies in setting up
the right kind of \dblenquote{semiotic dynamics}. In the present case, this dynamics should gradually coordinate
sensory experiences, prototypical views, individuals, 
and names, both within a single individual and across the
population. This led to the notion of a {\scshape semiotic network}, which was a significant conceptual advance 
in the methodology of conceiving and carrying out language game experiments.

@@

1. The robots need a {\scshape repertoire of actions}. So they need a sensori-motor system that allows them to carry out 
repeatedly and reliably an action. They can themselves observe the action both in the visual domain,  
looking at their own body or looking at their body in the mirror, or through proprioception, using sensors attached 
to the motors and joints on the body. There are many ways in which an action repertoire can be built up. One way 
is through motor babbling: The robot makes random movements and thus explores the sensori-motor 
space in order to find possible coherent actions. In the experiments reported here, the motor actions are acquired 
by kinesthetic learning: The experimenter moves parts of the body of the robot. These movements are recorded by the robot
and he can then play the movement back.

@@

In our group, Tony Belpaeme became the main Ph.D researcher for Colour Naming, finishing his thesis 
in 2005.\footnote{The first extensive paper on this research appeared as: \cite{Steels:2005}.}
This work was then continued by Joris Bleys, who managed to scale up from colour naming to colour description with 
a thesis defended in 2010.\footnote{This thesis is published as another volume in the present series as: \cite{Bleys:2014}}
The colour domain has attracted several other language game researchers, attempting to explain 
universal trends in colour categorisation.\footnote{Most notably: \cite{Puglisi:2008}, \cite{Baronchelli:2010}.}

@@

When the \textsc{qrio} humanoid robots became available, we decided to use this existing work for pushing forward the state of 
the art in colour language research in two ways:
\begin{itemize}
\item To use more realistic circumstances of object perception. Because 
robots are moving around in three dimensions and see the situation from different points of view, the reflection 
of light changes and is different for speaker and hearer (see \tabref{f:scene-example}). This difference did not 
play a role in earlier experiments but 
it certainly plays an important role in the stabilisation and choice of colour prototypes. 
\item To focus on colour descriptions beyond basic colour terms, i.e. descriptions like 
\textit{slightly blue}, \textit{blue-green}, \textit{bluish green}, \textit{very bright}, \textit{shiny yellow}, \textit{dark red}, \textit{sky blue}, 
\textit{very dark bluish red}, etc. Indeed, it turns out that 
basic colour terms are used only in about 10 to 15 \% of the occurrences of colour descriptions. The rest are 
more complex expressions. Complex colour descriptions require scaling up the semantics and introducing grammar, they 
also require the introduction of language strategies as a layer of abstraction 
above lexical and grammatical constructions. 
\end{itemize}

@@

A language strategy has two components: (i) a particular way of conceptualising reality (and learning the categories involved
in this conceptualisation) and (ii) a particular grammatical pattern that suggests to the listener what strategy is intended 
and provides the contents to execute the strategy. For example, in the phrase \textit{very green}, the combination of an adverb 
\textit{very} and a basic colour adjective \textit{green} suggests a \dblenquote{graded membership strategy} where not only a basic colour prototype
(\textit{green}) is introduced but also how close the sample is to this prototype (\textit{very}). I will now introduce first how 
the conceptual aspect of a language strategy is defined using a procedural semantics system called IRL and then how the 
linguistic aspect is defined, using Fluid Construction Grammar.

@@

To work out the procedural semantics hypothesis, we need to address the question what the primitives are of 
these semantic programs, and what computational formalism is most suited to combine them. Already in 2000, I reported
a representation system for computational semantics, called IRL.\footnote{
The earliest paper introducing IRL is: \cite{Steels:2000}. 
A more recent description of a new implementation is reported as: \cite{Spranger:2012}.}
IRL stands for Incremental Recruitment Language, for reasons that will become clear soon. \is{IRL}
IRL uses a constraint language approach. Constraint programming languages were pioneered in the  \is{Constraint Programming}
seventies by Borning, Steele, and others. It is a very active field with its own conferences, 
organisations and journals\footnote{see \url{http://www.informatik.uni-trier.de/ley/db/journals/constraints/}.} 
with wide applications in user interfaces, scheduling problems, design, games, etc.

@@

Constraint networks are an obvious choice for grounded (natural) language processing. \is{constraint network}
For example, to interpret the sentence \textit{the ball behind
the block}, it is possible that there are many blocks, many balls, but only one ball that is behind a block. So the 
unique referent of \textit{the ball}, \textit{the block} and of the total noun phrase only become clear when 
all constraints have been applied. Natural languages clearly do not supply control information but assume that 
the interpreter is smart enough to interpret constraints based on data flow. The other properties of constraint 
programming are also important. The constraint network planner can be used for conceptualising what to say and 
the constraint network matcher is highly useful in parsing an incomplete language fragment or in learning the 
meaning of a new grammatical construction.

@@

What is the nature of the constraints that we will need for the semantics of natural language? 
Take a phrase like \textit{the ball behind the block}. It invokes the following constraints, also called 
semantic operations: 
\begin{itemize}
\item \textit{the ball} evokes a constraint to use a prototype [\textsc{ball}] to find a referent in the current scene.
\item The same holds for \textit{the block} which uses another prototype [\textsc{block}]. 
\item The preposition \textit{behind} implies another constraint, namely to use the spatial relation [\textsc{behind}] to filter a 
set of objects (namely those found by using the [\textsc{ball}] prototype) and retain only those that satisfy the behind-relation 
to an object in another set. 
\end{itemize}

@@

So IRL is in fact a way to integrate the broad set of mechanisms used in 
pattern recognition (including neural networks) and AI (including machine learning). 
For example, an operation could implement a radial-basis-function network for doing categorisation and 
use the network to identify an object in a set. It could not only use the network but also adapt 
the weights in the network. Another operation could implement categorisation based 
on discrimination trees (as discussed in \chapref{chap:4}) and would not only do it, but also expand the discrimination trees 
when necessary. Other operations could implement set-theoretic operations, 
for the semantics of connectives like {\itshape and}, {\itshape or}, and {\itshape not}, by carrying out
{\scshape union}, {\scshape intersection} and {\scshape complement} operations over sets. There is no 
limit to the possible semantic operations, except that they have to terminate in finite time, possibly giving up when 
they cannot reach a valid result. IRL just provides a framework to define these operations and make them available 
to be recruited by the constraint network planner. IRL uses LISP as the host language, which means that all 
the computational machinery of LISP is available to implement semantic operations and associated learning algorithms.

@@

\enlargethispage{1\baselineskip}
1. The first strategy is the {\scshape Basic Colour Strategy} which implements a prototype-based approach to 
the basic colour categories. This approach is similar to the one used in the Proper Naming Game, except that 
only the three colour-dimensions are profiled (the opponent channels red-green, yellow-blue and the brightness channel). 
The strategy is similar to the one shown earlier. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour-lenient ?s3 ?s2 ?cc)
(select-most-activated ?t ?s3)
\end{verbatim}
When the speaker says \textit{green}, this network is used together with a specific binding of ?cc to the prototype for 
green (denoted as [green]). 
\begin{verbatim}
(bind colour-category ?cc [green])
\end{verbatim}
A graphical representation of this network is shown in \figref{fig:basic-strat}.

@@

2. The second strategy is the {\scshape Graded Membership Strategy} which categorises the topic first using 
a basic colour but then qualifies how well its colour fits with the prototype. For example, we can say \textit{greenish} or
\textit{very green}. Some languages (such as Tarahumara spoken in Northern Mexico) obligatory mark graded membership using 
suffixes. The core of the constraint network is defined as follows. A graphical representation of the network is shown
in \figref{fig:graded}. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour ?s3 ?s2 ?cc)
(filter-by-membership ?s4 ?s3 ?mc)
(select-most-activated ?t ?s4)
\end{verbatim}
This network grabs the objects in the context and binds them to ?s1 using the equal-to-context operation. 
The profile-colour-dimensions operation picks out the colour properties of the various objects in ?s1 resulting in 
a new set ?s2. The filter-by-colour operation computes for each object in the set ?s2
how well they match with the prototype bound to ?cc, constructing a new set ?s3. The filter-by-membership operation 
then computes how well 
each object in ?s3 satisfies the graded membership qualification bound to ?mc, 
constructing ?s4. And then the select-most-activated operation picks out the best candidate and binds it to ?t. 
Other data flows within the same network are possible, for example to find the bindings of ?mc and ?cc given 
a context and a referent ?t that needs to be identified.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/graded.pdf}}
\caption{\label{fig:graded}Constraint network implementing the Graded Membership Strategy. It builds further on the network for Basic Colour
Categorisation but adds a constraint to handle graded membership. }
\end{figure}
\enlargethispage{1\baselineskip}
The network is entirely general for all expressions that use graded membership and gets instantiated for a particular 
case by specifying the semantic entities involved using the bind operation. For example, Tarahumara has the word 
\textit{sita} for a particular colour prototype and \textit{kame} for one type of graded membership: 
\begin{verbatim}
(bind colour-category ?cc [sita])
(bind membership-category ?mc [kame]) 
\end{verbatim}

3. The third strategy is the {\scshape Category Combination Strategy} which compounds a category into two new ones, in 
expressions like \textit{greenish blue} or \textit{red orange}. This compounding is not a simple intersection. The set of objects  
that are \textit{greenish blue} is not found by taking all the ones that are green and all the ones that are blue and then 
the intersection of the two. Rather, it appears that the discretisation of the colour landscape is scaled with respect 
to the colour which appears as head of the phrase and then categorisation is done. This is achieved with the IRL network 
shown in \figref{fig:combi}. 
%http://www.workwithcolour.com/red-orange-colour-hue-range-01.htm

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/combi.pdf}}
\caption{\label{fig:combi}Constraint network implementing the Colour Combination strategy leading to expressions like \textit{greenish blue}. 
The filter-by-colour operation now has an extra argument, namely category-sets,  
and there is a draw-category-set-to-category operation that rescales the basic colour category set.}
\end{figure}

4. The next strategy is the {\scshape Basic Modification Strategy} which modifies a colour along the brightness dimension, as in 
\textit{light green}, \textit{dull yellow}, etc. It is implemented by projecting the modifying dimension onto the basic colour space 
whereby the colour being modified lies on the projecting vector and a new discretisation of the colour space can be 
performed. The network for this strategy is shown in \figref{fig:modi}. 
%http://www.workwithcolour.com/red-orange-colour-hue-range-01.htm

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/modi.pdf}}
\caption{\label{fig:modi}Constraint network for the basic modification strategy, for expressions like \textit{light green}.}
\end{figure}

@@

1. There are first {\scshape lexical constructions} that associate a word with the introduction of a semantic entity  
such as a colour category or a qualifier for indicating graded membership. So the meaning for the word \textit{yellow} 
introduces a \textsc{bind}-operation, as in: 
\begin{verbatim}
((bind colour-category ?c [yellow]))
\end{verbatim}
When the word \textit{yellow} is encountered, a unit for this word is created with this \textsc{bind}-operation as meaning and the 
word-string itself as the form. It is the unit called Yellow-unit in \figref{fig:syntactic-graded}. 
The construction also introduces a lexical category (part of speech) for the word (in this case the lexical category 
is colour-category) and specifies which variables in the meaning can be linked with other variables in other
units. The variables are typed. In this case there is one variable ?c which is of type colour-category.

@@

2. Next there are {\scshape functional constructions} that determine in which semantic operation the meaning of a word 
will play a role. It is necessary to have this layer above lexical categories because the same word/category can have many 
different functions. The choice is determined by the lexical category, and the syntactic and semantic context as 
defined in functional and phrasal constructions. For example, an adjective can appear both in an adjectival 
function (as in \textit{the red block}) and as a predicate (as in \textit{the block is red}) but it can also be coerced 
into a nominal function (as in \textit{red is a beautiful colour}). The same word can also belong to more than one 
lexical category, for example, \textit{walk} can be a verb (\textit{He walks home}) as well as a noun (\textit{I take a walk}). \is{functional 
constructions}

@@

3. Finally there are {\scshape phrasal constructions} that pull together the functional units in order to create a phrase. 
\is{phrasal construction}
The phrase usually has additional semantic operations as well. \figref{fig:syntactic-graded},left shows an example. 
The ColourPhrase-Unit is constructed on-top of the Colour-Unit. It adds three additional operations to the meaning: 
\begin{verbatim}
((equal-to-context ?s1)
 (profile-colour-dimensions ?s2 ?s1)
 (select-most-activated ?t ?s3))
\end{verbatim}
It links to the variables supplied by Colour-unit (namely ?s3 and ?s2) and the unit itself has a variable that can 
be linked further, namely ?t, which is the topic of the whole utterance. 
The constituent is called Colour-Phrase.

\figref{fig:syntactic-graded}, right shows another example, namely the structure for an instantiation of the 
graded-membership strategy for the Tarahumara utterance \textit{sita kame}. It is again a Colour-Phrase.
We see that there are now 4 levels of hierarchy. The ColourMembership-Unit pulls together a colour-category and 
a membership-category to form a new colour-category which then becomes a ColourPhrase-Unit again.
\subsection{Influence of embodiment}
% {\bfshape Influence of embodiment}\\

@@

I report here just briefly one additional experiment, relevant for understanding 
the impact of embodiment on language. The experiment compares three conditions (see \figref{fig:grounding-comparison}): 
(i) simulated perception (left column), (ii) shared grounded perception (middle column), meaning that both robots 
are given the same perception, and (iii) individual grounded 
perception (right column). The latter is the normal case when two robots are looking at the scene. 
It furthermore compares results for the Basic Naming Strategy (implemented with the IRL constraint network in 
\figref{fig:modi}) for four experimental conditions: 
\begin{enumerate}
\item There is a {\scshape baseline} experiment which starts out with the basic English colour categories provided 
by design to the agents. 
\item An {\scshape acquisition} experiment, where one agent (the tutor) is initialised with a set of basic colour categories and 
the other agent (the learning) has to acquire them. 
\item A {\scshape formation} experiment which involves a population of ten agents which develop a 
colour ontology and lexicon from scratch. 
\item And an {\scshape adaptation} experiment that starts with agents that are initialised with the 
English categories but then are allowed to further adapt their lexicons depending on the challenges of the situations 
they encounter. 
\end{enumerate}
